--- 
bibliography: references.bib
output:
  bookdown::pdf_book:
    toc: no
    template: null
    latex_engine: xelatex
    keep_tex: yes
    includes:
      in_header: parameters/preamble.tex
      before_body: parameters/doc_preface.tex
  bookdown::gitbook: default
  html_document:
    toc: yes
  bookdown::word_document2:
    fig_caption: yes
    md_extensions: +footnotes
    reference_docx: parameters/uwo_thesis_template.docx
    toc: yes
documentclass: book
classoption: oneside
fontsize: 12pt
subparagraph: yes
link-citations: no
biblio-style: apalike
always_allow_html: yes
---
```{r setup-index, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
options(digits = 3)
```

# Dissertating in RMarkdown + Bookdown: A preliminary guide

This tutorial was last updated: `r format(Sys.time(), '%d %B, %Y')`

This is a non-exhaustive guide to writing your dissertation using `RMarkdown` and `Bookdown`. 
Specifically, it will walk you through *one method* of organziing, writing, and rendering a dissertation with these tools, using an adapted version of the [Western University](uwo.ca) [thesis templates](https://grad.uwo.ca/academics/thesis/formatting.html).
This tutorial was written by me, [Thea Knowles](theaknowles.com). At the time of writing, I am currently in the throes of dissertating. This means that there are likely several details I haven't quite hammered out yet, or techniques I've missed. In the last year and a half, I've been collecting other people's tutorials and resources on using RMarkdown + for the purposes of using it to write a dissertation. The final product is my interpretation of these resources, adapted to my needs, and presented here as a *"What-I've-learned-so-far"*-style tutorial.

## Prerequisites

In order to use this tutorial, you need the following:

- [R](https://www.r-project.org/)
- [RStudio](https://www.rstudio.com/products/rstudio/download/)
  - Recent versions of RStudio also include [Pandoc](), which is required to compile documents
- Latex for [Mac](https://tug.org/mactex/mactex-download.html) or [Windows](https://miktex.org/download) (if you want to compile to PDF).
  - Alternatively, install [TinyTex](https://yihui.name/tinytex/), the Latex distribution created and recommended by Yihui Xie, creator of RMarkdown and bookdown[^139].

[^139]: TinyTex is probably the best way to go because Yihui always anticipates the problems we will run into, but I personally have not used it.

- R packages:

```{r eval=FALSE, echo=TRUE}
if(!require(devtools))
  install.packages("devtools", repos = "http://cran.rstudio.com")
install.packages("bookdown")
install.packages("knitr")

```

