---
title: "Figuring out figures with ggplot"
subtitle: "An R-Ladies #LdnOnt presentation"
author: "Thea Knowles"
date: "2018/10/25 (updated: `r Sys.Date()`)"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
    css: [default, metropolis, metropolis-fonts] 
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE, cache = TRUE)
library(emo)
library(plyr)
library(ggplot2)
```

# Plan for today

.pull-left[
**1. Philosophy of data visualization**

- Exploration vs. communication
- What is this *ggplot* and why do I care?

**2. Quick teaser of plotting possibilities**

- Because why not

**3. Workflow**

- RProjects
- Reading in data
- Looking at data
]

.pull-right[
**4. Making figures (finally!)**

- Basic set up
- Essential elements
- Popular plots
- Customization
- Unneccessary but delicious extra customizations as time allows

<img src = "images/ggplot_examples.jpeg">

]

---
# Get the data for today

## https://goo.gl/S7LFHC

---
# Philosophy of data visualization

## Exploration vs. Communication

### Exploration:
<center>
<iframe src="https://giphy.com/embed/l4KibOaou932EC7Dy" width="480" height="294" frameBorder="0" class="giphy-embed" allowFullScreen></iframe><p><a href="https://giphy.com/gifs/binoculars-l4KibOaou932EC7Dy">via GIPHY</a></p>
</center>

---
# Philosophy of data visualization

## Exploration vs. Communication

### Communication

<center>
<iframe src="https://giphy.com/embed/3o6Mb4y3SvguAcG3Ly" width="480" height="366" frameBorder="0" class="giphy-embed" allowFullScreen></iframe><p><a href="https://giphy.com/gifs/season-10-the-simpsons-10x22-3o6Mb4y3SvguAcG3Ly">via GIPHY</a></p>
</center>

---
# Philosophy of data visualization
## **gg**plot: **G**rammar of **G**raphics

.pull-left[
<img src = "images/gigi.jpg" height=500>
]

.pull-right[

### Essential elements of the "grammar":
- **Data**: Dataset to plot
- **Aesthetics**: Scales on to which data is mapped
- **Geometries**: visual elements used for the data
]


#### Goal of telling you all this:
- **Conceptualization**
- **Helpful terms to google** (mostly this)

---
class: middle

# ggplot: Grammar of Graphics

Plots are built up additively using *layers*. They usually look something like this:

```{r eval = FALSE}
ggplot(my_data, aes(x = my_x, y = my_y, color = my_group))+
     geom_someShape()+
     do_some_stuff()+
     do_other_stuff()
```

Each layer is added by a "+" sign
---

# Let's take a quick look at how this works

Throughout today's workshop, we will use a modified version of the [Bike Sharing Dataset](https://www.kaggle.com/marklvl/bike-sharing-dataset#Bike-Sharing-Dataset.zip) from [Kaggle](https://www.kaggle.com/). This dataset logs how many bikes are rented on a daily basis using the bike share system in Washington, D.C.

<center>
<img src = "images/bikeshare.jpg" height = 400>
</center>


```{r, echo = FALSE}
library(tidyverse)
bikes <- read.csv("data/bikes_new.csv")
```

---
# Let's take a quick look at how this works

## 1. Set up

.pull-left[
```{r plot-example1, eval = FALSE}
{{ggplot(bikes)}}
```
]

.pull-right[
```{r plot-example-out1, ref.label = "plot-example1", echo = FALSE}

```

]

---
# Let's take a quick look at how this works

## 2. Specify *data* and *aesthetics*

.pull-left[
```{r plot-example2, eval = FALSE}
{{ggplot(bikes, aes(x = temp, y = cnt))}}
```
]

.pull-right[
```{r plot-example-out2, ref.label = "plot-example2", echo = FALSE}

```

]

---
# Let's take a quick look at how this works

## 3. Specify the *geometry*

.pull-left[
```{r plot-example3, eval = FALSE}
ggplot(bikes, aes(x = temp, y = cnt)) +
     {{ geom_point() }}
```
]

.pull-right[
```{r plot-example-out3, ref.label = "plot-example3", echo = FALSE}

```
]

---
# Let's take a quick look at how this works

## 4. Specify *other stuff!*

.pull-left[
```{r plot-example4, eval = FALSE}
ggplot(bikes, aes(x = temp, y = cnt)) +
     geom_point()+
     {{geom_smooth(method = "lm")}}
```
]

.pull-right[
```{r plot-example-out4, ref.label = "plot-example4", echo = FALSE}

```
]


---
# Let's take a quick look at how this works

## 5. Prettify!

.pull-left[
```{r plot-example5, eval = FALSE}
ggplot(bikes, aes(x = temp, y = cnt)) +
     {{geom_point(aes(color = season))}}+
     geom_smooth(method = "lm")
```
]

.pull-right[
```{r plot-example-out5, ref.label = "plot-example5", echo = FALSE}

```
]

---
# Let's take a quick look at how this works
## 5. Prettify!

.pull-left[
```{r plot-example6, eval = FALSE}
ggplot(bikes, aes(x = temp, y = cnt)) +
     geom_point(aes(color = season))+
     geom_smooth(method = "lm")+
     {{theme_bw()}}
     
```
]

.pull-right[
```{r plot-example-out6, ref.label = "plot-example6", echo = FALSE}

```
]

---
# Let's take a quick look at how this works
## 5. Prettify!

.pull-left[
```{r plot-example7, eval = FALSE}
ggplot(bikes, aes(x = temp, y = cnt)) +
     geom_point(aes(color = season))+
     geom_smooth(method = "lm")+
     theme_bw()+
     {{facet_wrap(~season, scales = "free")}}
     # etc...
```
]

.pull-right[
```{r plot-example-out7, ref.label = "plot-example7", echo = FALSE}

```
]


---
# Workflow: R Projects

---
class: middle

# Workflow: R Projects

- [R Studio Projects](https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects) allow you to create *projects* in a given directory
- Benefits: 
     - Don't need to specify working directory
     - Everything is nice and neatly contained
     - Can open multiple projects at once (if you want)
     - Picks up where you left off (tabs open, environment, etc.) when you reopen the project

---
class: middle

# Exercise 1: R Projects

### Make an R Project

In the same directory as your .rproj file, you will include:
- the data bikes.csv
- a new .R script

---
class: center, middle

# Exercise 1: R Projects

### Make an R Project


### Step 1:
<img src = "images/rproj1.png">

---
class: center, middle

# Exercise 1: R Projects

### Make an R Project


### Step 2:
<img src = "images/rproj2.png">

---
class: center, middle

# Exercise 1: R Projects

### Make an R Project

### Step 3:
<img src = "images/rproj3.png">

---

# Exercise 1: R Projects

.pull-left[
### Make an RProject

In the same directory as your .rproj file, include:
- the data bikes.csv
- a new .R script entitled bikes.r
]

.pull-right[
### Step 4:
<img src = "images/rproj4.png">
]

*For now*: don't worry about *Creating a git repository* unless you are already familiar and set up with git. Git is extremely lovely and helpful and seamlessly integrated into R Studio, but that will be for another time.

---

# Exercise 1: R Projects

.pull-left[
### Make an RProject

In the same directory as your .rproj file, include:
- **the data bikes.csv**
     - Move the file to the directory you just created
- **a new .R script entitled bikes.r**
     - Open your new ggplot_workshop.rproj file and create a new R script and save it in the same directory as your project and data files.
]

.pull-right[
### Step 5:
<img src = "images/rproj5.png">


<img src = "images/rproj6.png">
]

---

# Workflow
## Loading and exploring data

---
# Excercise 2: Loading data

.pull-left[
**Option 1: Load in your data via the `Import dataset` option from the menu**

<img src = "images/importdata.png">

**Note**: the code used to execute this *will* appear in the console; you can also copy this code to your script for later (but keep in mind you don't actually *need* to specify the directory if you are using Projects)

]

--

.pull-right[
**Option 2. Load in your data in your script**
 
- This is (IMO) a better option because it makes it easy to do this everytime you open up the script (so, fewer things for you to click!)

In your new script, include the following lines:

```
# Read in your data
bikes <- read.csv("bikes_new.csv")
```

Anything preceded by `#` will be a *comment*, i.e., R will not try to run it (notes to self, etc.)
]

---

# Excercise 3: Exploring data

Let's see what each of these commands does...

```
View(bikes)
str(bikes)
summary(bikes)
head(bikes)
tail(bikes)
```

---
# Excercise 3: Exploring data

**Let's see what each of these commands does...**

.pull-left[
```{r eval = FALSE}
{{View(bikes)}}
str(bikes)
summary(bikes)
head(bikes)
tail(bikes)
```
]

.pull-right[
- Allows you to look at your data in the Source pane
- Looks and feels like Excel
]

---
# Excercise 3: Exploring data

**Let's see what each of these commands does...**

.pull-left[
```{r eval = FALSE}
View(bikes)
{{str(bikes)}}
summary(bikes)
head(bikes)
tail(bikes)
```
]

.pull-right[

- Gives you a breakdown of each variable (column):
     - Name, type of variable, data points (rows)
]

```{r echo = FALSE}
bikes <- read.csv("data/bikes_new.csv")
str(bikes)
```

---

# Excercise 3: Exploring data

**Let's see what each of these commands does...**

.pull-left[
```{r eval = FALSE}
View(bikes)
str(bikes)
{{summary(bikes)}}
head(bikes)
tail(bikes)
```
]

.pull-right[

- Gives you a summary of each variable:
     - **If numeric**: summary statistics (min/max, quartiles, mean)
     - **If factor**: levels of factors and n of observations in each level
]

```{r echo = FALSE}
summary(bikes)
```

---
# Excercise 3: Exploring data

**Let's see what each of these commands does...**

.pull-left[
```{r eval = FALSE}
View(bikes)
str(bikes)
summary(bikes)
{{head(bikes)}}
{{tail(bikes)}}
```
]

.pull-right[

- Shows you the first (head) or last (tail) 6 rows of your data
- **Note:** You can specify more or less than 6 rows (the default) like this: `head(bikes, 10)`
]

```{r echo = FALSE}
head(bikes)
```

---
class: middle

# A quick note on tidy data

- We won't get into multiple data formats today, but in general, "tidy data" is easy to work with for plotting (and everything else).
- **Tidy data** means that each row is a *single observation* and each column is a *single variable*

---
#LET'S PLOT SOME STUFF!

<center>
<iframe src="https://giphy.com/embed/AGOPaltgJ2pBC" width="402" height="480" frameBorder="0" class="giphy-embed" allowFullScreen></iframe><p><a href="https://giphy.com/gifs/thumbs-up-michelle-obama-cool-story-bro-AGOPaltgJ2pBC">via GIPHY</a></p>
</center>

---
class: middle

# Setting up a plot
## Load libraries

- In your ggplot_workshop.R script, add a line to the top to load in the libraries we will need to make our figures

```
library(ggplot2)
```

- You may need to install ggplot2 first. If you get an error saying *there is no package called 'ggplot2'*, run the following in your console first:

```
install.packages("ggplot2")
```
---
class: middle

# Plotting essentials
We will now iteratively build a figure together in ggplot

### Recall:
```{r eval = FALSE}
ggplot(my_data, 
       aes(x = my_x,
           y = my_y, 
           color = my_group))+
     geom_someShape()+
     do_some_stuff()+
     do_other_stuff()
```

---
# Plotting essentials

.pull-left[
## Setting up...

```{r plot-setup-1, eval = FALSE}
{{ggplot(bikes)}}
```
]

.pull-right[
```{r plot-setup-1-out, ref.label = "plot-setup-1", echo = FALSE}

```
]

---
# Plotting essentials: Aesthetics


.pull-left[
```{r plot-setup-2, eval = FALSE}
{{ggplot(bikes) + 
         aes(x = temp, 
             y = count)}}
```

- Recall from when we were exploring our data...      
     - `count`: number of bikes rented/day
     - `temp`: the temperature (F) that day
- We will now map those variables to our **aesthetics**
- Note what changes: there are no data yet because we haven't specified *how* those data will appear (that's the **geometry**)
]

.pull-right[
```{r plot-setup-2-out, ref.label = "plot-setup-2", echo = FALSE}

```

- Note the **+** at the end of the first line now - we are adding a new *layer*
]

---
# Plotting essentials: Geometry

.pull-left[
```{r plot-setup-3, eval = FALSE}
ggplot(bikes) +
       aes(x = temp, 
           y = count) +
     {{geom_point()}}
     
```

- We will now specify the **geometry** of our plot, i.e., how we want the data to be represented
- We will plot our data as points for now
- We'll look at more geometries soon (called *geoms* and specified as `geom_type()`)

]

.pull-right[
```{r plot-setup-3-out, ref.label = "plot-setup-3", echo = FALSE}

```

]

---
# Exploring other aesthetics

- Now that we've got a basic plot, we may wish to add other informative attributes to it.
- For example, we could map another variable to an aesthetic like `color`, `size`, `alpha`, etc...


<center>
<img src = "images/aesthetics.png">
</center>

See a comprehensive explanation of ggplot aesthetic specifications [here.](https://ggplot2.tidyverse.org/articles/ggplot2-specs.html)
---

# Exploring other aesthetics

- Now that we've got a basic plot, we may wish to add other informative attributes to it.
- For example, we could map another variable to an aesthetic like `color`, `size`, `alpha`, etc...
- Aesthetics like `color`, `size` (in general) get added to the main `ggplot()` function like this:

```{r eval = FALSE}
ggplot(bikes)+ 
       aes(x = temp, 
           y = count, 
           {{color = season}}, 
           {{size = holiday}}) +
     geom_point()
```

In the above example, we have mapped:

- `color` to the `season` variable (categorical variable with levels "fall", "winter", "summer", "spring")
- `size` to the `holiday` variable (categorical variable with levels 0 or 1, denoting whether that day was a holiday or not)

---
# Exploring other aesthetics
.pull-left[

```{r plot-setup-4, eval = FALSE}
ggplot(bikes, 
       aes(x = temp, 
           y = count, 
           {{color = season}})) +
     geom_point()
     
```

- Notice: within the `aes()` function, we are not adding *layers*; new aesthetics within the ( ) are just added by commas. This is still part of the basic `ggplot()` function
]


.pull-right[
```{r plot-setup-4-out, echo = FALSE}
ggplot(bikes, 
       aes(x = temp, 
           y = count, 
           color = season)) +
     geom_point()
```

]
--
**Exercise 4**: Adding aesthetics

Try modifying the current plot by mapping `alpha` (transparency) to `windspeed`

---

# Exercise 4: Adding aesthetics

Try modifying the current plot by mapping `alpha` (transparency) to `windspeed`

```{r, eval=FALSE}
ggplot(bikes, 
       aes(x = temp, 
           y = count, 
           color = season)) +
     geom_point()
     
```

---
# Exercise 4: Adding aesthetics

Try modifying the current plot by mapping `alpha` (transparency) to `windspeed`

.pull-left[
```{r plot-ex5, eval=FALSE}
ggplot(bikes, 
       aes(x = temp, 
           y = count, 
           color = season,
           {{alpha = windspeed}})) +
     geom_point()
     
```
]

.pull-right[
```{r plot-ex5-out, echo = FALSE}
ggplot(bikes, 
       aes(x = temp, 
           y = count, 
           color = season,
           alpha = windspeed)) +
     geom_point()
```
]
---
class: middle

# More customization of aesthetics
## Setting colours

- ggplot default colors are nice and pretty. BUT what if you want to define your own color scheme?

> A new layer: `scale_color_manual()`

- See also: `scale_color_continuous`, `scale_color_discrete`, and many more...

---
class: middle

# More customization of aesthetics
## Setting colors

### Name the colors
`scale_color_manual(values=c("red", "blue", "yellow", "green"))`

### Reference the [color hex codes](https://www.color-hex.com/)

`scale_color_manual(values=c("#d80e0e", "#1f0ed8",  "#ecee1e", "#11cc3a"))`

or

`my_colors <- c("#d80e0e", "#1f0ed8",  "#ecee1e", "#11cc3a")`
`scale_color_manual(values = my_colors)`

#### Use a predefined palate

`scale_color_brewer(palette="Set1")`

```
library(wesanderson)

plot + scale_color_manual(values = wes_palette(n = 4, name="Moonrise3"))`
```

```{r, include = FALSE}
# Set up ggplot default colors
gg_color_hue <- function(n) {
     hues = seq(15, 375, length = n + 1)
     hcl(h = hues, l = 65, c = 100)[1:n]
}
n = 4
cols = gg_color_hue(n)
```

---
class: middle
# More customization of aesthetics

- Note that we have mapped aesthetics to variables in these examples. You could also set aesthetics as static *attributes* (i.e., set the color of all points to *red* instead of to a variable).

- We have just looked at one tiny aspect of aesthetic changes. Check out the [ggplot documentation](https://ggplot2.tidyverse.org/articles/ggplot2-specs.html) for more. 

- Now we'll move on down from **aesthetics** to **geometries**
---

# Exploring geometries

- **Geometries**: How your data are represented, i.e., the *type* of plot (scatter, bar, line, etc.)

- The geometry you choose is important and should reflect the structure of the relationships in your data you want to show.

- There are [**lots**](https://ggplot2.tidyverse.org/reference/#section-layer-geoms). Here are some commonly used examples. We will walk through some of these. 

     - geom_point()
     - geom_line()
     - geom_bar()
     - geom_violin()
     - geom_boxplot()
     - geom_density()
     - geom_smooth()
     - geom_ridgeline()

---

# Geometries: Continuous x, y
.pull-left[
```{r point, eval = FALSE}
p <- ggplot(bikes) +
       aes(x = temp, 
           y = count)

p + 
     {{geom_point()}}
```
]

.pull-right[
```{r point-out, ref.label="point", echo = FALSE}

```
]

---
# Geometries: Continuous x, y

.pull-left[
```
p + 
     {{geom_smooth()}}
```
]

.pull-right[
```{r echo = FALSE}
p + geom_smooth()
```
]

---
# Geometries: Continuous x, y

**Combining geoms!**

.pull-left[
```{r, eval = FALSE}
p + 
     {{geom_point()}} +
     {{geom_smooth()}}
```
]
.pull-right[
```{r, echo = FALSE}
p + geom_point() +
     geom_smooth()
```
]

---
# Geometries: Categorical x, continuous y

.pull-left[
```{r eval = FALSE}
p <- ggplot(bikes) +
     aes(x = season,
         y = count)

p + 
     {{geom_boxplot()}}
```
]

.pull-right[
```{r echo = FALSE}
p <- ggplot(bikes) +
     aes(x = season,
         y = count)

p + geom_boxplot()
```
]
---
# Geometries: Categorical x, continuous y

.pull-left[
```
p + {{geom_violin()}}
```
]

.pull-right[
```{r echo = FALSE}
p + 
     geom_violin()
```

]
---
# Geometries: Categorical x, continuous y

.pull-left[
**Combining geoms!**
```{r, eval = FALSE}
p + 
     {{geom_violin()}} +
     {{geom_jitter(alpha = 0.25)}}

```

- `geom_jitter()` is basically `geom_point()` but the points get scattered (er, jittered) around to make them more visible
- here we have also specified the `alpha` aesthetic as a constant *attribute* within the geom.
]

.pull-right[
```{r, echo = FALSE}
p + 
     geom_violin() +
     geom_jitter(alpha = 0.25)

```
]

---
# Geometries: Histograms and density plots

.pull-left[
```{r eval = FALSE}
p <- ggplot(bikes) +
     aes(x = temp)

p + 
     {{geom_histogram()}}
```
]

.pull-right[
```{r echo = FALSE}
p <- ggplot(bikes) +
     aes(x = temp)

p + 
     geom_histogram()
```
]
---

# Geometries: Histograms and density plots

.pull-left[
```{r eval = FALSE}
p + 
     {{geom_freqpoly()}}
```
]

.pull-right[
```{r echo = FALSE}
p+geom_freqpoly()
```

]

---
# Geometries: Histograms and density plots

.pull-left[
```{r density-plot, eval = FALSE}
p + 
     {{geom_density()}}
```
]

.pull-right[
```{r density-plot-out, ref.label="density-plot", echo = FALSE}

```

]

---

# Geometries: Histograms and density plots

.pull-left[
```{r, eval = FALSE}
p <- ggplot(bikes) +
     aes(x = temp,
         {{fill = season}},
         {{alpha = 0.25}})
p + geom_density()
```
]

.pull-right[
```{r, echo = FALSE}
p <- ggplot(bikes) +
     aes(x = temp,
         fill = season,
         alpha = 0.25)
p + geom_density()
```
]

- What would happen if we specified `color` instead of `fill` for this geom?
---
# Geometries: Continuous x, categorical y

.pull-left[
```{r joyplot, eval = FALSE}

#install.packages("ggridges")
library(ggridges)
p <- ggplot(bikes)+ 
     aes(x = temp, 
         y = season)

p + geom_density_ridges()
```
]

.pull-right[
```{r joyplot-out, ref.label="joyplot", echo = FALSE}

```

]


---
# Summary plots


.pull-left[
We just saw plots of the raw data, but often we need to plot summary figures. To do that, we have to first manipulate our data

```{r}
bikes$season <- factor(bikes$season, levels = c("winter", "spring", "summer", "fall"))

bikes_summary <- bikes %>%
     ddply(.(season, type), 
           summarise, 
           count = mean(count))
```
]

.pull-right[
```{r}
bikes_summary
```

]

---
# Summary plots

.pull-left[
Now with our summarized data (`bikes_summary`)...

```{r bar, eval = FALSE}
ggplot(bikes_summary) +
     aes(x=season, y = count) +
     geom_bar(stat = "identity")
```
]

.pull-right[
```{r bar-out, ref.label="bar", echo = FALSE}

```

]
---
# Summary plots

.pull-left[
Now with our summarized data (`bikes_summary`)...

```{r bar-fill, eval = FALSE}
ggplot(bikes_summary) +
     aes(x=season, y = count, fill = type) +
     geom_bar(stat = "identity")
```
]

.pull-right[
```{r bar-fill-out, ref.label="bar-fill", echo = FALSE}

```

]

---
# Summary plots

.pull-left[
Now with our summarized data (`bikes_summary`)...

```{r line, eval = FALSE}
ggplot(bikes_summary) +
     aes(x=season, y = count, color = type) +
     geom_point()+
     {{geom_line(aes(group = type))}}
```
]

.pull-right[
```{r line-out, ref.label="line", echo = FALSE}

```

]

---
# Facets

- In addition to aesthetics, another way to distinguish different elements of our data is to separate them in panels.
- **Facets** allow us to do that.

.pull-left[
Now with our summarized data (`bikes_summary`)...

```{r bar-fill-facet, eval = FALSE}
ggplot(bikes_summary) +
     aes(x=season, y = count, fill = type) +
     geom_bar(stat = "identity") +
     {{facet_wrap(~type)}}
```

]

.pull-right[
```{r bar-fill-facet-out, ref.label="bar-fill-facet", echo = FALSE}

```

]

---
# Facets

- In addition to aesthetics, another way to distinguish different elements of our data is to separate them in panels.
- **Facets** allow us to do that.

.pull-left[
Now with our summarized data (`bikes_summary`)...

```{r bar-fill-facet2, eval = FALSE}
ggplot(bikes_summary) +
     aes(x=season, y = count, fill = type) +
     geom_bar(stat = "identity") +
     {{facet_grid(season~type, scales = "free")}}
```

]

.pull-right[
```{r bar-fill-facet2-out, ref.label="bar-fill-facet2", echo = FALSE}

```

]


---
# Prettifying your plot: Theme
## I.e., fonts and labels and backgrounds, oh my!


---
# Labels

.pull-left[
```{r labs-example, eval = FALSE}
p +
     {{labs(title = "Number of bikes rented per season",
          x = "Season",
          y = "Number of bikes rented",
          fill = "Type of bike rental")}}
```
]

.pull-right[
```{r labs-example-out, ref.label="labs-example", echo = FALSE}

```

]

---
# Exploring theme options

- Another important layer is the [`theme()`](https://ggplot2.tidyverse.org/reference/theme.html) layer, which allows you to control all non-data components of your plot.

> **Full disclosure:** 
- I always forget the syntax for the different elements of theme and look them up FREQUENTLY.
- I also keep an RMarkdown file called n00b.rmd that contains snippets of code I find myself looking up all the time, as a kind of *recipe-book* for myself

---
# Exploring theme options

- Helper functions are sometimes used to specify theme options. These begin with `element_`:
     - E.g., `element_text()` supports theme options to modify the text

- An example syntax: 

`theme(text.to.change = element_text(font = "font", size = 12, etc...))`

--

- You can set theme options on a plot-by-plot basis OR you can define custom themes with `theme_set()`
     - This is beyond the scope of today

--
**ALSO!**

- Check out [ggplotThemeAssist](https://github.com/calligross/ggthemeassist) and [ggedit](https://github.com/metrumresearchgroup/ggedit)
- I have never used these but they look super handy

```{r theme-example-setup, echo = FALSE}
p <- ggplot(bikes_summary) +
     aes(x = season, 
         y = count, 
         fill = type) +
     geom_bar(stat = "identity")
```

---
# Theme options: Some examples

.pull-left[
```{r theme-examples, eval = FALSE}
p + 
     {{theme(axis.title = element_text(size=18),
           axis.text.x = element_text(size=12),
           legend.title = element_text(size=12),
           legend.position = "bottom")}}
```
]

.pull-right[
```{r theme-examples-out, ref.label="theme-examples", echo = FALSE}

```

]

---
# Saving!

.pull-left[
## Option 1:

<img src = "images/exportplot.png">
]

--
.pull-right[
## Option 2: [ggsave](https://ggplot2.tidyverse.org/reference/ggsave.html)


```{r eval = FALSE}
my_plot <- ggplot(bikes_summary) +
     aes(x = season, 
         y = count, 
         fill = type) +
     geom_bar(stat = "identity")

ggsave("my_plot.png", width = 4)
```

- You do not need to specify the directory if you are using R Projects
- You could specify a relative directory (e.g., images/...)
]

---
# Extra fun stuff

### Random
- [accidental aRt:](https://twitter.com/accidental__art?lang=en) Dataviz gone beautifully wrong
- [data art](https://www.r-graph-gallery.com/portfolio/data-art/)
- [gganimate]()

### Colours
- [paletter package](http://djnavarro.net/post/2018-06-14-paletter/)
- [colorbrewer](http://colorbrewer2.org/#type=sequential&scheme=BuGn&n=3)

