---
title: "Starbucks data: Preliminary Results"
author: "Thea Knowles"
date: '2018-03-06'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r loadLibraries, include=FALSE}
# You don't have to include libraries you loaded in helper.R
# But you can if you want
# You may wish to load additional libraries as well

library(kableExtra)
```

```{r source, include=FALSE}
source('helper.R')
```

This is a summary report of the Starbucks drink menu data. This text will appear in the main output. 

<!-- This is an example of a comment I've written to myself that won't appear on the output -->
<!-- You can make a line a comment by selecting it and then typing shift + command + c -->
<!-- Alternatively, you can click Code >> Comment/Uncomment -->


Whenever we want to use r code, we embed it in chunks.

- Chunks can be inserted by typing alt + command + i
- Anything within a chunk will be treated like R code. 
- In chunks, you can either call variables that you assigned from within your helper script, or you can create new variables to include here. 


### Research question: 
- In Starbucks beverages, what is the relationship between the amount of sugars and calories?

### Means
- This is an example where we call in a variable we already created (calsugs, which is a data.frame)
- Note: what does `echo = FALSE` mean?

```{r means, echo=FALSE}
calsugs
```

#### Better looking means
- Be careful not to have two chunks with identical names!

```{r means_nice, echo=FALSE}
library(kableExtra)
knitr::kable(calsugs)

# You can include comments in chunks too, but you use the comment syntax that you normally would with r (#)

# If kableExtra() doesn't work for you, try using kable() instead
# kable() no longer functions with the most recent version of R
```


### A preliminary plot!
- You can also leave a chunk unnamed and without any parameters specified
- This is a plot that we made in helper.R

```{r}
calsugs_plot
```

### Another plot!
- Here we call in the same plot, but add some more details

```{r}
calsugs_plot+
     facet_wrap(~Beverage_category)
```

### OMG ANOTHER plot!
- Here we are making a plot from scratch 
     - i.e., we're not calling a plot that has already been assigned to a variable in helper.R
- We don't need to assign it to a variable name (<-); we can just include the code for the plot

**Calories by beverage type:**

```{r}
ggplot(data=starbucks, aes(x=Beverage_category, y=Calories, fill=Beverage_category))+
     geom_boxplot()
```


## Exercise 3: Knit!
- Press Knit button at top of RStudio editor
- Output will default to whatever is first specified in the YAML
     - In this case: HTML
- You can also specify other types of output and select which one you would like to knit

## Exercise 4: Include a new plot!
- Insert a chunk and make a new plot
- Use the same syntax that we used to make the "Calories by beverage type" but plot "Sugars by beverage type" instead


*Insert new chunk here!*


## Exercise 5: Make a new summary .Rmd report
- See slides



